{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-22 thesis/education.tex
bugs found at u chicago, and associated discussion

\noindent Landslide also includes several P2-specific instrumentations and features to cope with various student irregularities:
\begin{itemize}
	\item Quicksand emits different combinations of {\tt within\_function}/{\tt without\_function} directives
		for Landslide depending on the name of the test.
		For example, for {\tt paradise\_\allowbreak{}lost} Landslide will not preempt in a function named {\tt critical\_section()},
		which the test case uses to protect an internal counter used to detect the bug;
		and it will not preempt in any of the {\tt thr\_*()} thread library API functions
		for tests intended to target just the concurrency primitives.
		In future work this could be improved as annotations to be placed inside the test case code itself.
	\item Landslide finds ad-hoc synchronization patterns,
		such as {\tt while (!flag) yield()} or {\tt while (xchg(...)) continue},
		which students often open-code rather than using the prescribed synchronization API,
		and treats them as synchronization points as described in \sect{\ref{sec:landslide-blocking}}.
	\item Landslide finds ``too suspicious'' spinwait-loops in mutex implementations
		which are neither yield- nor {\tt xchg}-loops (as described above),
		which would ordinarily be classified as infinite loop bugs,
		and reports them with a suggestive message ({\tt undesirable\_\allowbreak{}loop\_html()} in {\tt landslide.c})
		referring the student to the appropriate lecture material
		\cite{synchronization-2}.
	\item The {\tt landslide} wrapper script logs the time and command-line options of invocation
		and captures a snapshot of the student code and results of the test and saves them to AFS after each run.
		% HURDLE_VIOLATION -- no, this is for p3 (context switcher) features only
\end{itemize}

@@

As described in \sect{\ref{sec:landslide-setup}},
all setup from the user's point of view is handled through the {\tt pintos-setup.sh} script.
It
and its helper {\tt pebsim/pintos/import-pintos.sh}
perform most of the same sanity checks as listed in \sect{\ref{sec:education-pebbles-instrumentation}},
then applies the patch {\tt annotate-\allowbreak{}pintos.patch}
(plus several more hacks in the script itself)
to insert the {\tt tell\_landslide()} annotations (\sect{\ref{sec:tell-landslide}})
into the student's kernel code.
The following tricks serve to make sure the annotations apply consistently
to (almost) all variations of code that students commonly submit:

\begin{itemize}
	\item Finds the declaration of {\tt ready\_list}, the scheduler runqueue declared by the basecode,
		and detects if the student has modified to be an array of lists rather than a single one.
		If so, defines the length of that array in a macro to be used by {\tt is\_runqueue()}
		(part of the patch described below).
		Either way defines a function {\tt get\_rq\_addr()} to return the address of the (first) list.
	\item Changes the basecode's definition of {\tt TIME\_SLICE} from 4 to 1 (units of timer ticks)
		so Landslide's timer injection will properly drive the context switcher.
	\item Inserts {\tt tell\_landslide\_forking()} into {\tt thread.c}
		(using {\tt sed} rather than the patch, described below,
		because it must go in a function which students have to implement,
		which is likely to disturb the context and make a patch fail).
	\item Adds the new {\tt priority-donate-multiple} test.
	\item Applies the {\tt annotate-pintos.patch} patch to the imported student implementation, which:
	\begin{itemize}
		\item Adds {\tt tell\_landslide\_thread\_on\_rq()}
			and {\tt tell\_landslide\_thread\_off\_rq()}
			annotations
			to {\tt list\_insert()} and {\tt list\_remove()} respectively
			(in {\tt lib/kernel/\allowbreak{}list.c}, which the students don't modify),
			which
			check whether the argument list
			is the scheduler runqueue
			using a helper function {\tt is\_runqueue},
			which in turn uses {\tt get\_rq\_addr()} and {\tt READY\_LIST\_LENGTH} described above.
		\item Modifies the existing {\tt priority-sema} and {\tt alarm-simultaneous} tests to be more Landslide-friendly.
		\item Inserts the {\tt tell\_landslide\_sched\_init\_done()},
			{\tt tell\_landslide\_vanishing()},
			and {\tt tell\_landslide\_thread\_switch()}
			annotations in the appropriate places
			(which the students generally do not modify).
	\end{itemize}
	\item Detects if the student has renamed the {\tt elem} field of the TCB struct,
		and if so renames its use in {\tt is\_runqueue()} (described above) correspondingly.
\end{itemize}

@@

\begin{itemize}
	\item {\tt priority-sema}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads to wait on a semaphore and signals them.
		Replaces threads with different priorities
		(originally chosen to produce deterministic output which the test checked for)
		with threads of the same priority.
	\item {\tt alarm-simultaneous}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads which each invoke {\tt timer\_sleep()} for a different amount of time.
		Number of (threads,iterations) reduced from (3,5) to (2,1).
	\item {\tt wait-simple}:
		Unmodified from the basecode's version, for {\em userprog}.
		Userspace process {\tt exec()}s a child process, which immediately exits, and {\tt wait()}s on it.
	\item {\tt priority-donate-multiple}:
		Written by Kevin Zhao, TA at U. Chicago, for {\em threads}.
		Tests for a priority donation race during {\tt lock\_release()}
		in which a thread holding a lock can accidentally keep a contending thread's donated priority
		after finishing releasing it.
\end{itemize}
\vspace{1em}

The (unpatched versions of) the first three tests are available at
\url{https://github.com/Berkeley-CS162/group0/tree/master/pintos/src/tests}.
The fourth test is available at
\url{https://github.com/bblum/landslide/blob/master/pebsim/pintos/priority-donate-multiple.c}.

@@

The data set is comprised of Landslide's automatically-generated usage snapshots
from CMU students from semesters Spring 2015 to Spring 2018 inclusive,
% and PSU \psuos from Spring 2018, % no snapshots from psu
CMU's official project grades from same,
CMU students' survey responses and submitted project code from Fall 2017 and Spring 2018
PSU students' survey responses and submitted project code from Spring 2018,
and U. Chicago students' survey responses from Fall 2017.
Table~\ref{tab:photo-of-ze-studence} shows the student participation rate across semesters
% TODO update for psu
for the thread library study.
% and PSU

@@

@@

\subsubsection{Response data}

@@

The response distributions for each of the surveys' multiple choice questions
are shown in Figure~\ref{fig:survey}.
In total, 28 students (or pairs thereof) answered the survey:
12 pairs from CMU, 15 individuals from PSU, and 1 pair from U. Chicago.
The first four questions/graphs focus on concrete debugging results,
and the latter six on the students' subjective opinions.
%
Note that two of the questions
(3 and 7 from \sect{\ref{sec:education-survey-pebbles}})
were not asked on U. Chicago's version of the survey,
so their corresponding graphs show only CMU and PSU response data.
Likewise, U. Chicago's questions 4, 8, and 10
(see \sect{\ref{sec:education-survey-pintos}}),
which were not asked on CMU's and PSU's surveys,
having only one respondent,
are not pictured;
the answers thereto were ``15-20 minutes'', ``Agree'', and ``Agree'', respectively.

@@

\subsubsection{False positives}

@@

		leading to undefined behaviour.

@@

The survey was introduced into the study only in time for two semesters' worth of student responses,
after several iterations of collecting only usage and bug report snapshots.
Apart from the obvious improvement of having been surveying students from the beginning,
the following questions could have improved the survey.
\begin{enumerate}
	\item {\em Did you have any technical difficulties with Landslide that I had to intervene on,
		whether in person or over email?}
		(Some students reported this in the ``false positives'' question,
		although fewer than I helped overall, so others must have not mentioned it.)
		Comparing answers to this question across subsequent semesters would give a sense
		of how much Landslide's stability was improving over time
		and whether it was mature enough for unsupervised use in the future.
	\item {\em How well do you feel you understood the research challenges explained in the lecture?}
		and,
		{\em How well do you feel a user should need to understand same in order to benefit from Landslide's bug reports?}
		(Answers on a scale from ``Not at all; Landslide is a magic black box to me''
		to ``I'm ready to work on research in this field myself''.)
		These questions would help fine-tune the lecture material and user guide to maximize student comfort,
		and potentially also corroborate the claim that Landslide is accessible even to novice users.
	\item {\em What additional debugging information would you want displayed on the preemption traces?}
		Knowing now that interpreting preemption traces was a sticking point for many users,
		I would hope to identify the most wished-for features to know what to prioritize improving.
		This could also assess students' understanding of what kinds of information
		would or would not be reasonable for Landslide to record and report.
\end{enumerate}

@@

Regarding non-research use in Pintos classes,
Landslide can now handle a considerably wider variety of student implementation quirks
on account of the fixes from this time (\sect{\ref{sec:education-pintos-instrumentation}}).
In its current shape I would recommend it for TA use grading,
but not necessarily directly to students without someone familiar with the codebase on immediate hand for tech support.
However, I also believe Landslide's success in these user studies,
provided me present to handle technical issues,
serves as testament for stateless model checking in general in the educational theatre.
While Pintos's kernel-level environment presents a unique challenge for concurrency testing,
other, more readily automatic model checkers for user-space programs,
such as dBug \cite{dbug-ssv},
could easily be used on other thread-library-like programming projects at any university.

>>>>

\noindent Landslide also includes several P2-specific instrumentations and features to cope with various student irregularities:
\begin{itemize}
	\item Quicksand emits different combinations of {\tt within\_function}/{\tt without\_function} directives
		for Landslide depending on the name of the test.
		For example, for {\tt paradise\_\allowbreak{}lost} Landslide will not preempt in a function named {\tt critical\_section()},
		which the test case uses to protect an internal counter used to detect the bug;
		and it will not preempt in any of the {\tt thr\_*()} thread library API functions
		for tests intended to target just the concurrency primitives.
		In future work this could be improved as annotations to be placed inside the test case code itself.
	\item Landslide finds ad-hoc synchronization patterns,
		such as {\tt while (!flag) yield()} or {\tt while (xchg(...)) continue},
		which students often open-code rather than using the prescribed synchronization API,
		and treats them as synchronization points as described in \sect{\ref{sec:landslide-blocking}}.
	\item Landslide finds ``too suspicious'' spinwait-loops in mutex implementations
		which are neither yield- nor {\tt xchg}-loops (as described above),
		which would ordinarily be classified as infinite loop bugs,
		and reports them with a suggestive message ({\tt undesirable\_\allowbreak{}loop\_html()} in {\tt landslide.c})
		referring the student to the appropriate lecture material
		\cite{synchronization-2}.
	\item The {\tt landslide} wrapper script logs the time and command-line options of invocation
		and captures a snapshot of the student code and results of the test and saves them to AFS
		(CMU's network file system)
		after each run.
		% HURDLE_VIOLATION -- no, this is for p3 (context switcher) features only
\end{itemize}

@@

As described in \sect{\ref{sec:landslide-setup}},
all setup from the user's point of view is handled through the {\tt pintos-setup.sh} script.
It
and its helper {\tt pebsim/pintos/import-pintos.sh}
perform most of the same sanity checks as listed in \sect{\ref{sec:education-pebbles-instrumentation}},
then applies the patch {\tt annotate-\allowbreak{}pintos.patch}
(plus several more hacks in the script itself)
to insert the {\tt tell\_landslide()} annotations (\sect{\ref{sec:tell-landslide}})
into the student's kernel code.
The following tricks were developed after trial-and-error on student submissions from the same semester,
and serve to make sure the annotations apply consistently
to (almost) all variations of commonly-submitted code.

\begin{itemize}
	\item Finds the declaration of {\tt ready\_list}, the scheduler runqueue declared by the basecode,
		and detects if the student has modified to be an array of lists rather than a single one.
		If so, defines the length of that array in a macro to be used by {\tt is\_runqueue()}
		(part of the patch described below).
		Either way defines a function {\tt get\_rq\_addr()} to return the address of the (first) list.
	\item Changes the basecode's definition of {\tt TIME\_SLICE} from 4 to 1 (units of timer ticks)
		so Landslide's timer injection will properly drive the context switcher.
	\item Inserts {\tt tell\_landslide\_forking()} into {\tt thread.c}
		(using {\tt sed} rather than the patch, described below,
		because it must go in a function which students have to implement,
		which is likely to disturb the context and make a patch fail).
	\item Adds the new {\tt priority-donate-multiple} test.
	\item Applies the {\tt annotate-pintos.patch} patch to the imported student implementation, which:
	\begin{itemize}
		\item Adds {\tt tell\_landslide\_thread\_on\_rq()}
			and {\tt tell\_landslide\_thread\_off\_rq()}
			annotations
			to {\tt list\_insert()} and {\tt list\_remove()} respectively
			(in {\tt lib/kernel/\allowbreak{}list.c}, which the students don't modify),
			which
			check whether the argument list
			is the scheduler runqueue
			using a helper function {\tt is\_runqueue},
			which in turn uses {\tt get\_rq\_addr()} and {\tt READY\_LIST\_LENGTH} described above.
		\item Modifies the existing {\tt priority-sema} and {\tt alarm-simultaneous} tests to be more Landslide-friendly.
		\item Inserts the {\tt tell\_landslide\_sched\_init\_done()},
			{\tt tell\_landslide\_vanishing()},
			and {\tt tell\_landslide\_thread\_switch()}
			annotations in the appropriate places
			(which the students generally do not modify).
	\end{itemize}
	\item Detects if the student has renamed the {\tt elem} field of the TCB struct,
		and if so renames its use in {\tt is\_runqueue()} (described above) correspondingly.
	\item Detects if the student has renamed the {\tt cur} (currently running thread) variable
		in the context switcher, and if so renames it back.
\end{itemize}

@@

\begin{itemize}
	\item {\tt priority-sema}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads to wait on a semaphore and signals them.
		Replaces threads with different priorities
		(originally chosen to produce deterministic output which the test checked for)
		with threads of the same priority.
	\item {\tt alarm-simultaneous}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads which each invoke {\tt timer\_sleep()} for a different amount of time.
		Number of (threads,iterations) reduced from (3,5) to (2,1).
	\item {\tt wait-simple}:
		Unmodified from the basecode's version, for {\em userprog}.
		Userspace process {\tt exec()}s a child process, which immediately exits, and {\tt wait()}s on it.
	\item {\tt wait-twice}:
		Unmodified from the basecode's version, for {\em userprog}.
		Slightly more complicated version of {\tt wait-simple},
		intended to expose failure-path bugs if the former finds no common-case ones.
	\item {\tt priority-donate-multiple}:
		Written by Kevin Zhao, TA at U. Chicago, for {\em threads}.
		Tests for a priority donation race during {\tt lock\_release()}
		in which a thread holding a lock can accidentally keep a contending thread's donated priority
		after finishing releasing it.
\end{itemize}
\vspace{1em}

The (unpatched versions of) the first four tests are available at
\url{https://github.com/Berkeley-CS162/group0/tree/master/pintos/src/tests}.
The fifth test is available at
\url{https://github.com/bblum/landslide/blob/master/pebsim/pintos/priority-donate-multiple.c}.

@@

The data set is comprised of Landslide's automatically-generated usage snapshots
from CMU students from semesters Spring 2015 to Spring 2018 inclusive,
% and PSU \psuos from Spring 2018, % no snapshots from psu
CMU's official project grades from same,
CMU students' survey responses and submitted project code from Fall 2017 and Spring 2018
PSU students' survey responses and submitted project code from Spring 2018,
and U. Chicago students' survey responses from Fall 2017.
Table~\ref{tab:photo-of-ze-studence} shows the student participation rate across semesters
% TODO update for psu
for the thread library study%
\footnote{
Participation at CMU was determined by receiving any automatic usage snapshots;
participation at PSU was determined, for lack of snapshots, by sign-up form submissions, which might over-count slightly.
}.
% and PSU

@@

Firstly, I sought to prove Landslide does as advertised:
finds concurrency bugs of severity, subtlety, and difficulty
consistent with the lessons an advanced operating systems class should hope to teach its students,
and provides diagnostic output that helps the student to understand and solve them.
% without necessarily spoiling the answer to the puzzle ... not scientifically evaluatable,
% but at least the trace format is de0u-approved.

At CMU, I configured Landslide to save a usage snapshot every time a student ran Landslide,
including command-line options issued, the current version of their project code,
a log of Landslide's output,
and the preemption traces for any bugs found.
These I analyzed by hand to determine how many distinct bugs were reported
(as multiple preemption traces may refer to the same bug; inspecting their code if necessary),
how many were deterministic
(i.e., Landslide reported them on the very first interleaving tested, with no need for artificial preemptions)
and how many were concurrency bugs,
whether any were false positives,
and whether the student was able to fix them thereafter
(determined if a subsequent snapshot showed them re-running the test and verifying the bug's absence;
an approximate measure at best, but far easier to implement than checking all the bugs by hand).

At U. Chicago,
course staff ran Landslide on student submissions behind-the-scenes
after each project deadline had passed,
and returned its preemption traces to any study volunteers.
I collaborated with the course staff
to confirm whether each trace represented a real bug and not a false positive
before distributing them to the students.
We decided on this procedure to optimize for most benefit to the students rather than for scientific rigor,
due to the small sample size (4 volunteering student groups among 21 submitted projects).

As PSU's version of the study was planned on relatively short notice,
and without immediate access to a shared, yet confidential, academic file system (such as CMU's AFS),
I was unable to examine its students' objective usage data or bug reports for this section.
Results from PSU are presented instead in \sect{\ref{sec:education-eval-survey}}.

\subsubsection{CMU}
\label{sec:education-eval-bugs-cmu}

% cmu - fps marked in the reuslts files? is this gonna be an accurate count?
\subsubsection{University of Chicago}
\label{sec:education-eval-bugs-uc}

Among the 4 groups volunteering to receive Landslide bug reports,
Landslide found 7 distinct bugs across the two projects tested,
which I confirmed by manual inspection of the students' code.
3 among these were deterministic; 4 were concurrency bugs.

\begin{enumerate}
	\item {\tt priority-donate-multiple} found 2 deterministic bugs and 1 concurrency bug in {\tt threads} projects.
		It found the targeted priority-leak bug
		described in \sect{\ref{sec:education-pintos-tests}} in one group,
		and exposed a {\tt NULL} pointer dereference in another
		when attempting to donate priority to an already-exited thread.
		The latter bug was observed deterministically.
		A third group neglected to implement priority donation during {\tt lock\_release()} at all,
		which was found deterministically.
	\item {\tt wait-simple} found 1 deterministic bug and 3 concurrency bugs in {\tt userprog} projects.
		It found a deadlock in two groups' implementations,
		% for one of these groups 2 traces were emitted, same stack trace (i think) just w diff pps
		wherein {\tt process\_wait()}'s synchronization assumed
		it would always be called before {\tt process\_exit()}.
		In both cases the former's {\tt cond\_wait()} call would block forever if reordered after the latter.
		This test also found several use-after-frees for one of those two groups,
		whose {\tt process\_exit()} didn't guard against a parent process exiting and deallocating its memory first;
		this root cause manifested in heap errors in 3 separate locations.
		Lastly, the heap checking alone found a deterministic use-after-free in a third group's {\tt exec()}.
\end{enumerate}

In cases where Landslide emitted multiple preemption traces for the same bug,
whether manifested the same way through different combinations of preemption points
or with different stack traces entirely arising from the same underlying cause,
course staff sent them all to the student,
including a disclaimer along the lines of ``some of these may indicate the same bug''.
%
The other 3 tests ({\tt priority-sema}, {\tt alarm-simultaneous}, and {\tt wait-twice}) found no bugs among the 4 groups.
One of the 4 groups had no bugs found among any of the 5 tests.

%%%% i think it'd be an irb violation even to publish this at all
%%%% and the potential false positive issue makes it not scientific enough anyway
% Course staff also ran Landslide with wait-simple on submissions from even non-volunteers;
% in total finding bugs in 11 of 18 groups' implementations.
% However, ... couldn't check for false positives because of non-volunteers ...

% XXX: did I actually fix this in the annotation scripts, or did I just hackhackhack to suppress it?
% my email to kevin just says, "I suppressed the false positive bug from wait-simple in GROUPNAME and got a real bug next".
Finally, one false positive was found while testing {\tt wait-simple}.
Landslide mistakenly reported that {\tt free()} had reentered {\tt malloc()},
which would be a potential heap corruption bug,
due to a technical discrepancy between when Pebbles and Pintos kernels take and release their heap allocation lock
with regard to the rest of the {\tt malloc()} API.
After discarding the bug report and suppressing this false positive,
Landslide moved on to find the true deterministic use-after-free described above.

Overall, I consider all of the 7 bugs to be ``severe'':
they are all either correctness violations (priority mis-donation)
or stability issues (crash, deadlock, or data corruption).
In the upcoming survey section, although only one of these groups did the survey,
they enthusiastically reported finding the bug reports very helpful (\sect{\ref{sec:education-survey-response-data}}).
Despite the small sample size, I consider this
a positive result that Landslide can be a useful concurrency programming aid
even at universities with different class projects than CMU's 15-410.
As for 1 false positive among 8 total reports,
it is difficult to decide what rate of mistakes is acceptable \cite{racerx},
but the upcoming survey results (\sect{\ref{sec:education-eval-survey-falsepositives}})
suggest that students are generally at least patient enough
to ask for help and make progress when technical support is available.

@@

\subsubsection{Response data}
\label{sec:education-survey-response-data}

@@

The response distributions for each of the surveys' multiple choice questions
are shown in Figure~\ref{fig:survey}.
In total, 28 students (or pairs thereof) answered the survey:
12 pairs from CMU, 15 individuals from PSU, and 1 pair from U. Chicago.
The first four questions/graphs focus on concrete debugging results,
and the latter six on the students' subjective opinions.
%
Note that two of the questions
(3 and 7 from \sect{\ref{sec:education-survey-pebbles}})
were not asked on U. Chicago's version of the survey,
so their corresponding graphs show only CMU and PSU response data.
Likewise, U. Chicago's questions 4, 8, and 10
(see \sect{\ref{sec:education-survey-pintos}}),
which were not asked on CMU's and PSU's surveys,
having only one respondent,
are not pictured;
the answers thereto were ``15-20 minutes'', ``Agree'', and ``Agree'', respectively.
Also, this respondent's answer of 6 on the ``How many bugs'' question
appears to indicate the total number of preemption trace files course staff sent them;
%despite the attached ``some of these may indicate the same bug'' disclaimer;
upon further investigation,
the 6 traces seem to represent 2 distinct bugs among them (\sect{\ref{sec:education-eval-bugs-uc}}).

@@

\subsubsection{False positives}
\label{sec:education-eval-survey-falsepositives}

@@

		scrambling its instrumentation and leading to chaotic behaviour.

@@

The survey was introduced into the study only in time for two semesters' worth of student responses,
after several iterations of collecting only usage and bug report snapshots.
Apart from the obvious improvement of having been surveying students from the beginning,
the following questions could have improved the survey.
\begin{enumerate}
	\item {\em Did you have any technical difficulties with Landslide that I had to intervene on,
		whether in person or over email?}
		(Some students reported this in the ``false positives'' question,
		although fewer than I helped overall, so others must have not mentioned it.)
		Comparing answers to this question across subsequent semesters would give a sense
		of how much Landslide's stability was improving over time
		and whether it was mature enough for unsupervised use in the future.
	\item {\em How well do you feel you understood the research challenges explained in the lecture?}
		and,
		{\em How well do you feel a user should need to understand same in order to benefit from Landslide's bug reports?}
		(Answers on a scale from ``Not at all; Landslide is a magic black box to me''
		to ``I'm ready to work on research in this field myself''.)
		These questions would help fine-tune the lecture material and user guide to maximize student comfort,
		and potentially also corroborate the claim that Landslide is accessible even to novice users.
	\item {\em What additional debugging information would you want displayed on the preemption traces?}
		Knowing now that interpreting preemption traces was a sticking point for many users,
		I would hope to identify the most wished-for features to know what to prioritize improving.
		This could also assess students' understanding of what kinds of information
		would or would not be reasonable for Landslide to record and report.
	\item {\em For each bug Landslide found in your code, how trivial or severe do you feel it was?}
		This would help get a sense of how the students regarded Landslide
		on a spectrum between annoying style check or as a life-saver,
		and potentially suggest options to make Landslide suppress certain types of bug reports.
		For example, it currently reports spin-wait loops in {\tt mutex\_lock()} as bugs
		with a special message referring students to relevant lecture slides,
		but it's possible refusing to test any code beyond until that bug was fixed
		might have made Landslide less useful overall.
\end{enumerate}

@@

Regarding non-research use in Pintos classes,
Landslide can now handle a considerably wider variety of student implementation quirks
on account of the fixes from this time (\sect{\ref{sec:education-pintos-instrumentation}}).
In its original shape
(before the F'17 semester,
having only enough instrumentation necessary for the Pintoses used in \sect{\ref{sec:quicksand-eval}}'s experiments),
Landslide was already able to automatically instrument 18 out of the 21 {\tt threads} project submissions at U. Chicago.
I was able to quickly deploy a fix to make the annotation scripts handle the other 3 cases,
although such technical support is not something any TA would be able to do.

In its current shape I would recommend it for TA use grading,
but not necessarily directly to students without someone familiar with the codebase on immediate hand for tech support.
However, I also believe Landslide's success in these user studies,
provided me present to handle technical issues,
serves as testament for stateless model checking in general in the educational theatre.
While Pintos's kernel-level environment presents a unique challenge for concurrency testing,
other, more readily automatic model checkers for user-space programs,
such as dBug \cite{dbug-ssv},
could easily be used on other thread-library-like programming projects at any university.
